# Topics in Exploratory Data Analysis I　{#topics1}

Prof. Taisei Kaizoji and R Notebook Compiled by HS


* Get AGCO2rv.csv from Moodle and place it at the same place as this file.

* Reference: Data Analysis for Researchers Week 7(02/02)
  - Inference Statistics Regression Analysis
  
### Summary {-}

1. Give a descriptive name to each variable, when importing data by `WDI` command. 
1. Save data of investigation by `write_csv`, but if you have RNotebook, you do not need to save at every step.
2. If you use pipe in `tidyverse`, i.e, %>%, properly, you do not need to `attach(data)` or `dettach(data)`.

### Dot Place Holder in `lm`. {-}

In the following, the following format is used.

```
df %>% lm(y ~ x, .)
```

This is same as 

```
lm(y ~ x, data = df)
```

'.' is called the dot place holder. Recall that most of the function (or command) we use in tidyverse, the first argument is the data frame. When we use piping, we omit the first argiment. For example,

```
df %>% ggplot(aes(x, y)) + geom_point()
```
 is same as
 
```
ggplot(df, aes(x,y)) + geom_point()
```

However, the data frame argument of `lm()` is not the first one, therefore we use the dot place holder to show the place the data frame should be inserted.


## Part I: Regression 1: 

Run the following only once, or you can install these packages using install packages in the Tool menu.

```
install.packages("car")
install.packages("modelsummary")
```


### Why does GDP vary from country to country?

```{r}
library(tidyverse) # tidyverse Package, a collection of packages for data science
library(WDI) # WDI Package for World Development Indicators
library(car) #VIF Tool for checking multi-collinearity
library(modelsummary) # Tool for writting tables of the regression results
```

* WDI Indicators
  - NY.GDP.MKTP.CD: GDP (current US$)
  - SP.POP.TOTL: Population, total
  - AG.LND.TOTL.K2: Land area (sq. km)

```{r cash = TRUE}
wb <- as_tibble(WDI(country="all", 
                    indicator=c(gdp = "NY.GDP.MKTP.CD", pop = "SP.POP.TOTL", land = "AG.LND.TOTL.K2"),
                    start=1960, end=2020,
                    extra=TRUE))
wb
```

<!-- ### Save wb as csv file. -->

<!-- ```{r} -->
<!-- write_csv(wb, file="wb_gdp_pop_land.csv") -->
<!-- ``` -->

```{r cash = TRUE}
wb_ag20 <- wb %>% filter(year == 2020 & region=="Aggregates")
wb_ag20
```


### Regression

```{r cash = TRUE}
wb_ag20 %>% ggplot(aes(pop, gdp)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "GDP and Population", 
       x = "total population", y = "GDP")
```


```{r}
gdp_pop <-wb_ag20 %>% lm(gdp ~ pop, .)
summary(gdp_pop)
```

```{r}
wb_ag20 %>% ggplot(aes(land, gdp)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "GDP and Land Area", 
       x = "land area", y = "GDP")
```


```{r}
gdp_land <-wb_ag20 %>% lm(gdp ~ land, .)
summary(gdp_land)
```

```{r}
gdp_pop_and_land <-wb_ag20 %>% lm(gdp ~ pop + land, .)
summary(gdp_pop_and_land)
```

```{r}
gpa_models <- list(gdp_pop = gdp_pop, gdp_land = gdp_land, gdp_pop_and_land = gdp_pop_and_land)
msummary(gpa_models, statistic = 'p.value')
```
* The default is: `msummary(gpa_models)`. `msummary(gpa_models, statistic = 'p.value')` replaces the standard error by p-value. Compare the following with the one above and the `Std.Error` of the summaries of each model.

```{r}
msummary(gpa_models)
```


```{r}
wb_ag20 %>% ggplot(aes(pop, land)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "population and Land", 
       x = "population", y = "land")
```


### Checking for multicollinearity 

```{r}
pop_land <-wb_ag20 %>% lm(pop ~ land, .)
summary(pop_land)
```



### vif: Determination of multi-collinearity

VIF = 1/(1-{R<sub>i</sub>}<sup>2</sup>)


#### VIF (Variance Inflation Factor)

```{r}
gdp_pop_and_land %>% vif()
```

#### Conclusion

* Regions which have large population have large GDP.
* It was also found that the larger the region, the more populated the region. This suggests that the GPs of regions with large regions and large populations are significant.
* This result suggests a positive relationship between population and economic activity.

## A time series data of World from 1960 to 2020


```{r}
wb_world <- wb %>% filter(country=="World") %>% arrange(year)
wb_world
```

```{r}
wb_world %>% ggplot() + 
  geom_line(aes(x = year, y = gdp)) +
  labs(title = "GDP (current US$)")
```

```{r}
wb_world %>% ggplot() + 
  geom_line(aes(x = year, y = pop)) +
  labs(title = "Total population)")
```

```{r}
wb_world %>% lm(gdp ~ pop, .) %>% summary()
```

```{r}
wb_world %>% ggplot() +
  geom_point(aes(pop, gdp)) +
  labs(title = "World GDP and World population", 
       x = "Total population", y = "GDP (current US$)")
```
```{r}
wb_world_extra <- wb_world %>% mutate(diff_gdp = gdp - lag(gdp), diff_pop = pop - lag(pop)) 
wb_world_extra
```

```{r}
wb_world_extra %>% ggplot(aes(x = year, y = diff_gdp)) + geom_line() +
  labs(title = "World GDP changes", y = "gdp changes")
```
```{r}
wb_world_extra %>% ggplot(aes(x = year, y = diff_pop)) + geom_line() +
  labs(title = "World population changes", y = "population changes")
```
```{r}
wb_world_extra %>% ggplot(aes(x = diff_pop, y = diff_gdp)) + geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "World GDP changes and World GDP changes", x = "population changes", y = "GDP changes") 
```
```{r}
wb_world_extra %>% lm(diff_gdp ~ diff_pop, .) %>% summary()
```

### Excercise: 

1.  Select a region from the following; 

  - East Asia & Pacific
  - Europe & Central Asia
  - Middle East & North Africa
  - Sub-Saharan Africa
  - Latin America & Caribbean

2. Preform regression analysis. 

3. Select the best model.

## Regression 2 : CO2 Emission

### Project

* Problem: “What are the causes of increased CO2 emissions?”
* Plan: To develop and test the hypothesis that CO2 emissions depends on GDP, population and land under cereal production of each region.
  
* Data: Using the World Bank's World Development Indicators (WDI), we will collect data on GDP, population and land area for regions around the world.
  - WDIindicators
  - GDP(Y):"NY.GDP.MKTP.CD"
  - Total Population (X1): "SP.POP.TOTL"
  - Land under cereal production (hectares): "AG.LND.CREL.HA"
  
#### Regression Analysis

* Regression equation:Y =c + a x1 + b x2 + d x3 Y: CO2 emission: independent variable
  - x1: GDP (Y) : independent variable
  - x2: Total Population (X1) : independent variable x3: Land under cereal production (hectares):
independent variable
* We perform a regression analysis of the above regression equation (estimate the parameters, c (intercept), a, b, d) and elect the best model.


#### WDI Indicator cordes

* co2 = EN.ATM.CO2E.KT: CO2 Emission 
* gdp =  NY.GDP.MKTP.CD:  GDP (current USD)
* pop = SP.POP.TOTL: Total Population
* cereal = AG.LND.CREL.HA:  Land under cereal production (hectares)

```{r}
WDIsearch(string = "EN.ATM.CO2E.KT", field = "indicator", short = FALSE)[[3]]
```


```{r cash = TRUE}
wb_co2<- as_tibble(WDI(country="all", 
                   indicator=c(co2 = "EN.ATM.CO2E.KT",
                               gdp = "NY.GDP.MKTP.CD",
                               pop = "SP.POP.TOTL",
                               cereal = "AG.LND.CREL.HA"),
                   start=1960,
                   end=2020,
                   extra=TRUE))
wb_co2
```


<!-- ### Save CO2 as csv file. -->

<!-- ```{r} -->
<!-- write_csv(wb_co2, file="wb_co2.csv") -->
<!-- ``` -->

In the lecture the folloiwng is used. However, regions are overlapping and it seems to be better to chooose non-aggregated data.

```{r}
wb_co2_16 <- wb_co2 %>% filter(year == 2016 & region=="Aggregates")
wb_co2_16
```

```{r}
wb_co2 %>% filter(region != "Aggregates", !is.na(co2)) %>% 
  group_by(year) %>% summarize(n = n_distinct(country)) %>% arrange(desc(n), desc(year))
```

Let us choose non-aggregated data in 2018.

```{r}
wb_co2_18 <- wb_co2 %>% filter(region != "Aggregates", !is.na(co2), year == 2018) 
wb_co2_18
```





### Regression

```{r}
wb_co2_18 %>% ggplot(aes(gdp, co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "CO2 emission and GDP")
```

```{r}
wb_co2_18 %>% ggplot(aes(gdp, co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  scale_x_log10() + scale_y_log10() +
  labs(title = "CO2 emission and GDP in Log-Log scale")
```
Let us take log-log plot

```{r}
wb_co2_18 %>% ggplot(aes(pop, co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  scale_x_log10() + scale_y_log10() +
  labs(title = "CO2 emission and Population in Log-Log scale")
```
```{r}
wb_co2_18 %>% ggplot(aes(cereal, co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  scale_x_log10() + scale_y_log10() +
  labs(title = "CO2 emission and Land under cereal production in Log-Log scale")
```


```{r}
wb_co2_18m <- wb_co2_18 %>% filter(!is.na(co2), !is.na(gdp), !is.na(pop), !is.na(cereal), cereal > 0)
co2_gdp <- wb_co2_18m %>% lm(log10(co2) ~ log10(gdp), .)
co2_pop <- wb_co2_18m %>% lm(log10(co2) ~ log10(pop), .)
co2_cereal <- wb_co2_18m %>% lm(log10(co2) ~ log10(cereal), .)
co2_gdp_pop <- wb_co2_18m %>% lm(log10(co2) ~ log10(gdp) + log10(pop), .)
co2_gdp_cereal <- wb_co2_18m %>% lm(log10(co2) ~ log10(gdp) + log10(cereal), .)
co2_pop_cereal <- wb_co2_18m %>% lm(log10(co2) ~ log10(pop) + log10(cereal), .)
co2_all <- wb_co2_18m %>% lm(log10(co2) ~ log10(gdp) + log10(pop) + log10(cereal), .)
```

```{r}
msummary(list(gdp = co2_gdp, pop = co2_pop, cereal = co2_cereal, gdp_pop = co2_gdp_pop, gdp_cereal = co2_gdp_cereal, pop_cereal = co2_pop_cereal, all = co2_all))
```
```{r}
vif(co2_gdp_pop)
vif(co2_gdp_cereal)
vif(co2_pop_cereal)
vif(co2_all)
```
```{r}
list(gdp_pop = vif(co2_gdp_pop), gdp_cereal = vif(co2_gdp_cereal), pop_cereal = vif(co2_pop_cereal), all = vif(co2_all))
```


#### We select regression model; 


CO2 emission = c + a (GDP) + b (population)

(Omit land because of multicollinearity)


#### A time series data of World from 1960 to 2020


```{r}
wb_co2_world <- wb_co2 %>% filter(country=="World")
wb_co2_world 
```


```{r}
wb_co2_world  %>% ggplot() + 
  geom_line(aes(x = year, y = co2)) +
  labs(title = "Co2 Emmision of the World")
```
```{r}
wb_co2_world %>% ggplot() + 
  geom_line(aes(x = year, y = gdp)) +
  labs(title = "GDP (current US$)")
```

```{r}
wb_co2_world_extra <- wb_co2_world %>% 
  mutate(diff_co2 = co2 - lag(co2), diff_gdp = gdp - lag(gdp))
wb_co2_world_extra
```

```{r}
wb_co2_world_extra %>% ggplot(aes(x = year, y = diff_co2)) + geom_line() +
  labs(title = "Changes of World CO2 emission")
```
```{r}
wb_co2_world_extra %>% ggplot(aes(x = year, y = diff_gdp)) + geom_line() +
  labs(title = "Changes of World GDP")
```

```{r}
wb_co2_world_extra %>% lm(diff_co2 ~ diff_gdp, .) %>% summary()
```


## Regression 3: CO2 emission (Standardization)

* Prof. Kaizoji provided a data in csv format in Moodle and use scale function to standardize the data, let us proceed one step by one step.
* He starts with a WDI data in 2016 and use the aggregated part. There may be arguments on the set of data, let us use the same one.
* He defines variance in his slides, and the codes below uses unbiased variance. The difference is minor.


```{r}
 wb_co2
```

```{r}
co2_ag16 <- wb_co2 %>% filter(year == 2016, region=="Aggregates") %>%
  select(iso2c, co2, gdp, pop, cereal) %>% drop_na()
co2_ag16
```

```{r}
co2_al16_std <- co2_ag16 %>% mutate(st_co2 = (co2 - mean(co2))/sd(co2),
                                    st_gdp = (gdp - mean(gdp))/sd(gdp),
                                    st_pop = (pop - mean(pop))/sd(pop),
                                    st_cereal = (cereal - mean(cereal))/sd(cereal))
co2_al16_std
```

```{r}
co2_al16_scaled <- co2_ag16 %>% mutate(st_co2 = scale(co2),
                                    st_gdp = scale(gdp),
                                    st_pop = scale(pop),
                                    st_cereal = scale(cereal))
co2_al16_scaled 
```

```{r}
co2_ag16 %>% select(-1) %>% scale() %>% as_tibble()
```


```{r}
co2_ag16_rv <- read_csv("data/AGCO2rv.csv") 
co2_ag16_rv
```

```{r}
colnames(co2_ag16_rv) <- c("st_co2", "st_gdp", "st_pop", "st_cereal")
co2_ag16_rv
```




### Regression

```{r}
co2_al16_scaled  %>% ggplot(aes(st_gdp, st_co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "CO2 emission and GDP: Standardized")
```

```{r}
co2_al16_scaled  %>% ggplot(aes(st_pop, st_co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "CO2 emission and Population: Standardized")
```
```{r}
co2_al16_scaled  %>% ggplot(aes(st_cereal, st_co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "CO2 emission and Land under cereal production: Standardized")
```


```{r}
st_co2_gdp <- co2_al16_scaled  %>% lm(st_co2 ~ st_gdp, .)
st_co2_pop <- co2_al16_scaled  %>% lm(st_co2 ~ st_pop, .)
st_co2_cereal <- co2_al16_scaled  %>% lm(st_co2 ~ st_cereal, .)
st_co2_gdp_pop <- co2_al16_scaled %>% lm(st_co2 ~ st_gdp + st_pop, .)
st_co2_gdp_cereal <- co2_al16_scaled  %>% lm(st_co2 ~ st_gdp + st_cereal, .)
st_co2_pop_cereal <- co2_al16_scaled  %>% lm(st_co2 ~ st_pop + st_cereal, .)
st_co2_all <- co2_al16_scaled  %>% lm(st_co2 ~ st_gdp + st_pop + st_cereal, .)
```

```{r}
msummary(list(st_gdp = st_co2_gdp, st_pop = st_co2_pop, st_cereal = st_co2_cereal, st_gdp_pop = st_co2_gdp_pop, st_gdp_cereal = st_co2_gdp_cereal, st_pop_cereal = st_co2_pop_cereal, st_all = st_co2_all))
```
```{r}
vif(st_co2_gdp_pop)
vif(st_co2_gdp_cereal)
vif(st_co2_pop_cereal)
vif(st_co2_all)
```


### Not Standardized

```{r}
co2_al16 <- wb_co2 %>% filter(region != "Aggregates", year == 2016)
co2_al16
```

### Regression

```{r}
co2_al16 %>% ggplot(aes(gdp, co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "CO2 emission and GDP")
```

```{r}
co2_al16 %>% ggplot(aes(pop, co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "CO2 emission and Population")
```
```{r}
co2_al16 %>% ggplot(aes(cereal, co2)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "CO2 emission and Land under cereal production")
```


```{r}
orig_co2_gdp <- wb_co2_18 %>% lm(co2 ~ gdp, .)
orig_co2_pop <- wb_co2_18 %>% lm(co2 ~ pop, .)
orig_co2_cereal <- wb_co2_18 %>% lm(co2 ~ cereal, .)
orig_co2_gdp_pop <- wb_co2_18 %>% lm(co2 ~ gdp + pop, .)
orig_co2_gdp_cereal <- wb_co2_18 %>% lm(co2 ~ gdp + cereal, .)
orig_co2_pop_cereal <- wb_co2_18 %>% lm(co2 ~ pop + cereal, .)
orig_co2_all <- wb_co2_18 %>% lm(co2 ~ gdp + pop + cereal, .)
```

```{r}
msummary(list(orog_gdp = orig_co2_gdp, orig_pop = orig_co2_pop, orig_cereal = orig_co2_cereal, orig_gdp_pop = orig_co2_gdp_pop, orig_gdp_cereal = orig_co2_gdp_cereal, orig_pop_cereal = orig_co2_pop_cereal, orig_all = orig_co2_all))
```
```{r}
vif(orig_co2_gdp_pop)
vif(orig_co2_gdp_cereal)
vif(orig_co2_pop_cereal)
vif(orig_co2_all)
```

#### Standadized (Again)

```{r}
msummary(list(st_gdp = st_co2_gdp, st_pop = st_co2_pop, st_cereal = st_co2_cereal, st_gdp_pop = st_co2_gdp_pop, st_gdp_cereal = st_co2_gdp_cereal, st_pop_cereal = st_co2_pop_cereal, st_all = st_co2_all))
```
```{r}
vif(st_co2_gdp_pop)
vif(st_co2_gdp_cereal)
vif(st_co2_pop_cereal)
vif(st_co2_all)
```


### We select regression model; 

* CO2 emission = c + a (GDP) + b (population)

* (Omit land because of multicollinearity)



## Regression of Categorical Variables and ANOVA

If you start from here, you need to load the following packages.

```{r}
library(tidyverse) #tidyverse Package, a collection of packages for data science
library(WDI) #WDI Package for World Development Indicators
library(car) #VIF function
library(modelsummary) #Table of the regression results
```

```{r}
wb_co2
```

```{r}
wb_co2_4anova <- wb_co2 %>% filter(year == 2016, region != "Aggregates") %>%
  select(iso2c, co2, gdp, income) %>% drop_na() %>%
  mutate(log_co2 = log10(co2), log_gdp = log10(gdp))
wb_co2_4anova
```

```
CO2 <- as_tibble(read.csv("CO2CAPITA.csv"))
CO2
str(CO2)
head(CO2)
tail(CO2)
```

```{r}
wb_co2_4anova %>% group_by(income) %>% 
  summarize(n = n_distinct(iso2c), co2_mean = mean(co2), gdp_mean = mean(gdp), log_co2_mean = mean(log_co2), log_gdp_mean = mean(log_gdp))
```



### Regression

```
y <- CO2$EN.ATM.CO2E.PC #CO2 Emission per capita
x <- CO2$NY.GDP.PCAP.PP.CD/10000 #GDP per capita, PPP
summary(x)
summary(y)
```

```{r}
wb_co2_4anova %>% ggplot(aes(log_gdp, log_co2)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "CO2 emission per capita and GDP per capita in log-log scale", 
       x = "GDP per capita", y = "CO2 emission per capita")
```

```
CO2 %>% ggplot(aes(x, y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "CO2 emission per capita and GDP per capita", 
       x = "GDP per capita", y = "CO2 emission per capita")

model_1 <-lm(y ~ x, data=CO2)
summary(model_1)
msummary(model_1, statistic = 'p.value')
```

```{r}
wb_co2_4anova %>% ggplot(aes(x = log_co2, fill =  income)) + 
  geom_histogram( ) 
```
```{r}
wb_co2_4anova %>% ggplot(aes(x = log_co2, y = income, fill =  income)) + 
  geom_boxplot()
```

```{r}
wb_co2_anova <- wb_co2_4anova %>% mutate(hi = as.numeric(income == "High income"))
```

```{r}
wb_co2_model <- wb_co2_anova %>% lm(co2 ~ gdp + hi, .)
wb_co2_model %>% summary()
```
```{r}
wb_co2_log_model <- wb_co2_anova %>% lm(log_co2 ~ log_gdp + hi, .)
wb_co2_log_model %>% summary()
```
```{r}
income_model <- wb_co2_anova %>% lm(co2 ~ factor(income), .) 
income_model %>% summary()
```


```{r}
msummary(list(co2_model = wb_co2_model, co2_log_model = wb_co2_log_model, co2_income_model = income_model), statistic = 'p.value')
```

#### t-test

```{r}
high <- wb_co2_4anova %>% filter(income == "High income") %>% pull(co2)
low <- wb_co2_4anova %>% filter(income == "Low income") %>% pull(co2)
t.test(high, low)
```

### Regression analysis with categorical variables

model_3=lm(EN.ATM.CO2E.PC ~ factor(income), data = CO2) 

summary(model_3) # summary

msummary(model_3, statistic = 'p.value')



### ANOVA (Analysis of Variance)

```{r}
wb_co2_4anova %>% aov(co2 ~ factor(income), .) %>% summary()
```


## Week 7 Assignment

* Week 7: Assignment on regression analysis and PPDAC cycle

* Work on the following problem.

* Problem: Find one problem from the SGDs goal.

* Example: WDI environment

* https://datatopics.worldbank.org/world-development-indicators/themes/environment.html

* Plan: Analyze the data to find out why the problem is occurring.

* Data: Collect data from WDI and other sources (OECD, UN, IMF, etc.).

* Analysis: Analyze the data. Use what you have learned about regression analysis and descriptive statistics in the previous lectures.

* Conclusion: Briefly describe what you learn from the data analysis.

* Submission: Your rmd file (Rmarkdown) on Moodle box. 

*Deadline : Feb. 8. 


